\vspace{-1in} Demystifying the Computer Science \\PhD Admission in US Universities\\{\large A Guide for International Students

### 2023-11-02 demystify.tex
updates

By around March you should hear back from most PhD programs you applied to . If you haven't heard back, reach out via email and ask about the status of your application.
If you receive offers, congratulations!  Now you're at a different game because the schools that have admitted you will try to get you to accept them!  Several important factors to consider are the reputation of schools and professors (\S\ref{sec:schoolsandprofs}), and of course, funding (\S\ref{sec:funding}). You will have to make your decision by around April 15.

\paragraph{Open House} Most schools will have \textbf{Open House} events, which are a great resource to learn about the school, department, faculty, research, living, etc. During the Open House, you get a chance to talk to individual faculty and current students.  Take notes of faculty who make you excited, count those that are taking in new students (if they meet you, likely they are considering new students!).  Talk to students about their advisors, the dept, the area, funding situation etc.  Ask about anything you want to determine that they deserve \emph{you}.
In summary, if possible, attend the Open House.  Even if you can't come in person, you still can attend virtually and meet with individual faculty.

\begin{commentbox}[Vu:]
GMU has \emph{Virtual} Open House (VOH), e.g., \url{https://cs-gmu.github.io/cs-phd-voh-s23/}, which I've co-organized in the last two years. We invite all admitted PhD students to the VOH through Zoom to learn about the CS program, the department, GMU, and the DC area in general. Students also get opportunities to chat with professors and current students.
\end{commentbox}

@@

\emph{Project proposal}: in many countries, you have to choose a project and advisor \emph{during} the application process (e.g., you write a proposal to a potential advisor). In the US, you often start your PhD without a project and advisor and find them later. This is also why US PhD programs are longer.

\emph{Course work}: in the US you will spend the first couple of years taking classes and explore potential advisor and research topics. In other countries, you often do not do course work and start your research right away, e.g., you immediately work on the research project you propose with the advisor you chose. In the US you also have to pass a series of "exams", e.g., qualifying exam, comprehensive exam, thesis proposal defense.  In other countries, you do not have to do these exams or only do few of them. For example, ABD (all but dissertation) is used in the US to refer to a PhD candidate who have finished all course work and exams and only need to write and defend their dissertation.

>>>>

By around March you should hear back from most PhD programs you applied to. If you haven't heard back, reach out via email and ask about the status of your application.
If you receive offers, congratulations!  Now you're at a different game because the schools that have admitted you will try to get you to accept them!  Important factors to consider are the reputation of schools and professors (\S\ref{sec:schoolsandprofs}), and of course, funding availability (\S\ref{sec:funding}). You will have to make your decision by around \emph{April 15}.

\paragraph{Open House} Most schools will have \textbf{Open House} events, which are a great resource to learn about the school, department, faculty, research, living, etc. Even if you can't come in person, you still can attend virtually and meet with individual faculty.
During the Open House, you get a chance to talk to individual faculty and current students.  Take notes of faculty who make you excited, count those that are taking in new students (if they meet you, likely they are considering new students!).  Talk to students about their advisors, the dept, the area, funding situation etc.  Ask about anything you want to determine that they deserve \emph{you}.

\begin{commentbox}[Vu:]
GMU has \emph{Virtual} Open House (VOH), e.g., \url{https://cs-gmu.github.io/cs-phd-voh-s23/}, which I've co-organized in recent years. We invite all admitted PhD students to the VOH through Zoom to learn about the CS program, the department, GMU, and the DC area in general. Students also get opportunities to chat with professors and current students.
\end{commentbox}

@@

\emph{Project proposal}: in many countries, you have to choose a project and advisor \emph{during} the application process (e.g., you write a proposal to a potential advisor). In the US, you often start your PhD without an advisor or project and find them later. This is also why US PhD programs are longer.

\emph{Course work}: in the US you will spend the first couple of years taking classes and explore potential advisor and research topics. In other countries, you often do not do course work and start your research right away, e.g., you immediately work on the research project you proposed with the advisor you chose. Also, in the US you also have to pass a series of "exams", e.g., qualifying exam, comprehensive exam, thesis proposal defense\footnote{The word "ABD" (all but dissertation) is used in the US to refer to a PhD candidate who have finished all course work and exams and only need to write and defend their dissertation.}. In other countries, you do not have to do these exams or only do few of them.
